---
title: "data_prep"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get HexIDs for High Seas

- [h3](https://crazycapivara.github.io/h3-r/articles/h3.html)

- [bbnj::p_abnj](https://benioffoceaninitiative.github.io/bbnj/reference/p_abnj.html)
- https://github.com/crazycapivara/h3-r/blob/master/inst/examples/nc-polyfill.R

- [h3:polyfill](https://crazycapivara.github.io/h3-r/reference/polyfill.html)\
  Get all hexagons with centers contained in a given polygon\
  `hex_ids <- polyfill(bbnj::p_abnj)`
  https://crazycapivara.github.io/h3-r/reference/polyfill.html 
  
- [h3::h3_set_to_multi_polygon](https://crazycapivara.github.io/h3-r/reference/h3_set_to_multi_polygon.html)\
  Create a (multi) polygon describing the outline(s) of a set of hexagons
  `hex_sf <- h3::h3_set_to_multi_polygon(hex_ids)`
  analyze existing raster layers with these to populate planning units and features for prioritizr

```{r libraries}
# remotes::install_github("crazycapivara/h3-r")
librarian::shelf(
  # deckgl, geojsonsf, ggplot2, h3jsr, jsonlite, rnaturalearth, 
  BenioffOceanInitiative/bbnj, crazycapivara/h3, 
  dplyr, glue, here, leaflet, mapboxapi, mapview, purrr, readr, rgdal, sf, units)
mapviewOptions(fgb = FALSE)
```   

## make `hex_sf` for `abnj` at different resolutions

```{r, eval=F}
# get polygons for ABNJ (outside EEZ)
abnj <- bbnj::p_abnj %>% 
 st_set_crs(4326) # %>% 
 # st_wrap_dateline()

# world
wrld <- bind_rows(
  st_sf(
    hemisphere = "west", 
    geom = st_as_sfc(
      st_bbox(
        c(xmin = -180, xmax = 0, ymin = -86, ymax = 90), 
        crs = st_crs(4326)))),
  st_sf(
    hemisphere = "east", 
    geom = st_as_sfc(
      st_bbox(
        c(xmin = 0, xmax = 180, ymin = -86, ymax = 90), 
        crs = st_crs(4326)))),
  st_sf(
    hemisphere = "global", 
    geom = st_as_sfc(
      st_bbox(
        c(xmin = -180, xmax = 180, ymin = -90, ymax = 90), 
        crs = st_crs(4326)))),
  st_sf(
    hemisphere = "test", 
    geom = st_as_sfc(
      st_bbox(
        c(xmin = -180, xmax = -160, ymin = 40, ymax = 90), 
        crs = st_crs(4326)))))

fix_dateline <- function(geom){
  # geom = g
  xy <- st_coordinates(geom)
  x <- xy[,'X']
  
  if (max(abs(diff(x))) < 300)
    return(geom)
  
  y <- xy[,'Y']

  if (x[1] < 0){
    x[x > 0] <- x[x > 0] - 360    
  } else {
    # browser("TODO:if (x[1] > 0)")
    x[x < 0] <- x[x < 0] + 360
  }

  z <- try(suppressWarnings(
    tibble(
      x = x,
      y = y) %>% 
      st_as_sf(coords = c("x", "y"), crs = 4326) %>%
      summarise(geometry = st_combine(geometry)) %>%
      st_cast("POLYGON") %>% 
      st_cast("MULTIPOLYGON") %>% 
      pull(geometry) %>% 
      # st_make_valid() %>% 
      st_wrap_dateline() %>%
      # st_make_valid() %>% 
      st_union() # %>% 
    # st_make_valid() %>% 
  ))
  # mapview(z)
  
  if ("try-error" %in% class(z))
    browser()

  z
}

make_hex_res <- function(hex_res = 1){
  # hex_res = 2
  
  # paths
  hex     <- glue("abnj_hex_res{hex_res}")
  hex_geo <- glue(here("data/{hex}.geojson"))
  # hex_shp <- glue(here("data/{hex}_shp/{hex}.shp"))
  # dir.create(dirname(hex_shp), showWarnings = F)
  
  # get hexagon ids for whole world (have to do by hemisphere)
  hexids <- c(
    h3::polyfill(wrld[1,], res = hex_res),
    h3::polyfill(wrld[2,], res = hex_res))
  # test area
  # hexids <- h3::polyfill(wrld[4,], res = hex_res)
  
  # convert hexagon ids to spatial features
  hex_sf <- map_df(hexids, h3_to_geo_boundary_sf) %>% 
    mutate(
      hexid = hexids)
  
  # fix hexagons crossing dateline
  hex_sf <- hex_sf %>% 
    mutate(
      on_dtln = map_lgl(geometry, function(g){ 
        max(diff(st_coordinates(g)[,'X'])) > 300 } )) %>% 
    rowwise() %>% 
    mutate(
      geometry = fix_dateline(geometry))
  # mapview(abnj) + mapview(hex_sf)

  # get intersection with Areas Beyond National Jurisdiction
  x      <- st_intersects(hex_sf, abnj, sparse = F)[,1]
  hex_sf <- hex_sf[x,]
  # mapview(abnj) + mapview(hex_sf)
  
  # add centroid lon, lat
 hex_sf <- hex_sf %>% 
   mutate(
    geometry = st_cast(geometry, "MULTIPOLYGON"),
    # ctr = map(geometry, st_centroid),
    lon = map_dbl(geometry, function(g){
      st_centroid(g) %>% st_coordinates() %>% 
        .[,'X'] }),
    lat = map_dbl(geometry, function(g){
      st_centroid(g) %>% st_coordinates() %>%
        .[,'Y'] })) # %>% 
 # mapview(abnj) + mapview(hex_sf)

 # clip hexagons to abnj
 # hex_sf_0 <- hex_sf
 # hex_sf <- hex_sf_0
 hex_sf <- hex_sf %>% 
   # st_wrap_dateline() %>% 
   st_intersection(
     # st_wrap_dateline(
       abnj
       # )
       ) %>% 
   filter(
     st_geometry_type(geometry) %in% c("POLYGON", "MULTIPOLYGON")) %>% 
   mutate(
     geometry = st_cast(geometry, "MULTIPOLYGON")) %>% 
   st_wrap_dateline()
 
 # st_geometry_type(hex_sf$geometry) %>% table()
 # mapview(abnj) + mapview(hex_sf)
 # leaflet(
 #  hex_sf, 
 #  options = leafletOptions(worldCopyJump = T)) %>% 
 #  addProviderTiles(providers$Esri.OceanBasemap) %>% 
 #  addPolygons()
 
  # write out geojson and shapefile outputs
  write_sf(hex_sf, hex_geo, delete_dsn=T)
  # write_sf(hex_sf, hex_shp, delete_dsn=T)
  return(T)
}

make_hex_res(1)
# sapply(1:5, make_hex_res)
sapply(3:5, make_hex_res)
```

```{r}
d <- tibble(
  hex_res = 1:4) %>% 
  mutate(
    h            = map(hex_res, function(r){
      # r <- 1
      g <- glue(here("data/abnj_hex_res{r}.geojson"))
      message(glue("reading {basename(g)}"))
      h <- read_sf(g) # mapview(h)
      h }),
    n_pu         = map_dbl(h, nrow),
    avg_area_km2 = map_dbl(h, function(h){
      # h <- d0$h[[3]]
      h %>% 
        mutate(geometry = st_make_valid(geometry)) %>% 
        # mutate(geometry = st_wrap_dateline(geometry)) %>% 
        # pull(geometry) %>% st_is_valid() %>% table()
        # pull(geometry) %>% st_geometry_type() %>% table()
        filter(st_is_valid(geometry)) %>%  # mapview(h) # missig some
        pull(geometry) %>% 
        # st_geometry_type() %>% table()
        st_area() %>%
        set_units(km^2) %>% 
        mean()  }),
    avg_width_km = map_dbl(avg_area_km2, sqrt))
d %>% 
  select(-h) %>% 
  write_csv(here("data/abnj_hex_res_summary.csv"))
```

TODO: investigate geometry repair over sf::`st_make_valid()` with wdpar::`st_repair_geometry`, especially for dateline issues:

* [Clean data — wdpa_clean • wdpar](https://prioritizr.github.io/wdpar/reference/wdpa_clean.html)
* [Repair geometry — st_repair_geometry • wdpar](https://prioritizr.github.io/wdpar/reference/st_repair_geometry.html)

```{r mbtiles, eval=FALSE}
abnj <- bbnj::p_abnj %>% 
  sf::st_set_crs(4326) # %>% 
  # st_wrap_dateline()

hex_res <- 2
hex_geo <- glue(here("data/abnj_hex_res{hex_res}.geojson"))
hex_sf <- read_sf(hex_geo)
# sf::st_geometry_type(hex_sf$geometry) %>% table()

leaflet(
  hex_sf, 
  options = leafletOptions(worldCopyJump = T)) %>% 
  addProviderTiles(providers$Esri.OceanBasemap) %>% 
  addPolygons()
mapview(hex_sf)

hex_res <- 2
hex_geo <- glue(here("data/abnj_hex_res{hex_res}.geojson"))
hex_mbt <- glue(here("data/abnj_hex_res{hex_res}.mbtiles"))

library(mapboxapi)

mapbox_token_txt <- '~/My Drive/private/mapbox_token_bdbest.txt'
Sys.setenv("MAPBOX_API_TOKEN" = readLines(mapbox_token_txt))

Sys.setenv(
  "PATH" = paste0(Sys.getenv('PATH'),":/Users/bbest/homebrew/bin"))

stopifnot(file.exists(hex_geo))

tippecanoe2 <- function (input, output, layer_name = NULL, min_zoom = NULL, 
    max_zoom = NULL, drop_rate = NULL, overwrite = TRUE, other_options = NULL, 
    keep_geojson = FALSE) 
{
    check_install <- system("tippecanoe -v") == 0
    if (!check_install) {
        stop("tippecanoe is not installed.  Please visit https://github.com/mapbox/tippecanoe for installation instructions.", 
            call. = FALSE)
    }
    opts <- c()
    if (!is.null(min_zoom)) {
        opts <- c(opts, sprintf("-Z%s", min_zoom))
    }
    if (!is.null(max_zoom)) {
        opts <- c(opts, sprintf("-z%s", max_zoom))
    }
    if (is.null(min_zoom) && is.null(max_zoom)) {
        opts <- c(opts, "-zg")
    }
    if (!is.null(drop_rate)) {
        opts <- c(opts, sprintf("-r%s", drop_rate))
    }
    else {
        opts <- c(opts, "-as")
    }
    if (overwrite) {
        opts <- c(opts, "-f")
    }
    collapsed_opts <- paste0(opts, collapse = " ")
    if (!is.null(other_options)) {
        extra_opts <- paste0(other_options, collapse = " ")
        collapsed_opts <- paste(collapsed_opts, extra_opts)
    }
    dir <- getwd()
    if (any(grepl("^sf", class(input)))) {
        input <- sf::st_transform(input, 4326)
        if (is.null(layer_name)) {
            layer_name <- stringi::stri_rand_strings(1, 6)
        }
        if (keep_geojson) {
            outfile <- paste0(layer_name, ".geojson")
            path <- file.path(dir, outfile)
            sf::st_write(input, path, quiet = TRUE, delete_dsn = TRUE, 
                delete_layer = TRUE)
        }
        else {
            tmp <- tempdir()
            tempfile <- paste0(layer_name, ".geojson")
            path <- file.path(tmp, tempfile)
            sf::st_write(input, path, quiet = TRUE, delete_dsn = TRUE, 
                delete_layer = TRUE)
        }
        call <- sprintf("tippecanoe -o %s/%s %s %s", dir, output, 
            collapsed_opts, path)
        browser("first")
        system(call)
    }
    else if (inherits(input, "character")) {
        if (!is.null(layer_name)) {
            collapsed_opts <- paste0(collapsed_opts, " -l ", 
                layer_name)
        }
        call <- sprintf("tippecanoe -o %s/%s %s %s", dir, output, 
            collapsed_opts, input)
        browser("last")
        # tippecanoe -o /Users/bbest/Github/BenioffOceanInitiative/bbnj-app//Users/bbest/Github/BenioffOceanInitiative/bbnj-app/data/abnj_hex_res2_x.mbtiles -Z0 -z2 -as -f -l hex_res2 /Users/bbest/Github/BenioffOceanInitiative/bbnj-app/data/abnj_hex_res2_x.geojson
        # tippecanoe -o /Users/bbest/Github/BenioffOceanInitiative/bbnj-app/data/abnj_hex_res2.mbtiles -Z0 -z2 -as -f -l hex_res2 /Users/bbest/Github/BenioffOceanInitiative/bbnj-app/data/abnj_hex_res2.geojson
        system(call)
    }
}

tippecanoe2(
  input = hex_geo, output = hex_mbt, layer_name = "hex_res2",
  min_zoom=0, max_zoom=2)

mb_access_token(
  readLines("~/My Drive/private/mapbox_bdbest_secret_token.txt"),
  overwrite = T, install = T)
# readRenviron("~/.Renviron")
Sys.getenv("MAPBOX_SECRET_TOKEN")

upload_tiles(
  input        = hex_mbt, 
  username     = "bdbest", 
  tileset_id   = "bbnj_hex_res2",
  tileset_name = "bbnj_hex_res2")
```


## map `hex_sf` with `leaflet`

```{r}
hex_res <- 4
hex_geo <- glue(here("data/abnj_hex_res{hex_res}.geojson"))
hex <- read_sf(hex_geo) # 129,991 × 3

mapview(hex)

# works with leaflet , though maybe CRS is off?
leaflet(data = hex, width = "100%") %>% 
  addProviderTiles("Stamen.Toner") %>% # providers$Esri.OceanBasemap
  addPolygons(
    # weight = 2,
    color = "white",
    fillColor = "white",
    fillOpacity = 0.7)
```

## map with plotly and mapbox vector tiles

`bdbest.9wxfwo42`

```{r}
library(plotly)

mapbox_token_txt <- '~/My Drive/private/mapbox_token_bdbest.txt'
mapbox_token     <-  readLines(mapbox_token_txt)
#Sys.setenv("MAPBOX_TOKEN" = readLines(mapbox_token_txt)) # for Orca

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    marker = list(color = "fuchsia"),
    type = 'scattermapbox',
    hovertext = us_cities[,"City"]) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = 'dark',
      zoom =2.5,
      center = list(lon = -88, lat = 34))) 
fig <- fig %>%
  # config(mapboxAccessToken = Sys.getenv("MAPBOX_TOKEN"))
  config(mapboxAccessToken = mapbox_token)

fig
```

## `mapboxer`

### get started

```{r}
library(mapboxer)

map <- motor_vehicle_collisions_nyc %>%
  dplyr::mutate(
    color = ifelse(injured == 0, "yellow", "red")
  ) %>%
  as_mapbox_source() %>%
  mapboxer(
    center = c(-73.9165, 40.7114),
    zoom = 10
  ) %>%
  add_navigation_control() %>%
  add_circle_layer(
    circle_color = c("get", "color"),
    circle_blur = 1,
    circle_stroke_color = "red",
    circle_stroke_width = 1,
    popup = "<p>{{date}} {{time}}</p><p>Number of persons injured: {{injured}}</p>"
  )

if (interactive()) map
```


### v

```{r}
library(mapboxer)

mapbox_token_txt <- '~/My Drive/private/mapbox_token_bdbest.txt'
Sys.setenv("MAPBOX_API_TOKEN" = readLines(mapbox_token_txt))

# SRC_ID <- "country-boundaries-data"
# 
# vector_src <- mapbox_source(
#   type = "vector",
#   url = "mapbox://mapbox.country-boundaries-v1")

# Style URL: mapbox://styles/bdbest/cl2139ehh000814piexs7is7f
# Access token: pk.eyJ1IjoiYmRiZXN0IiwiYSI6ImNqeTQzcjMzYjE2MnMzb3AyaXBmOWZueG4ifQ.kOxbgvSal1U-3LaYKOo3rw
# Integration URL: https://api.mapbox.com/styles/v1/bdbest/cl2139ehh000814piexs7is7f/wmts?access_token=pk.eyJ1IjoiYmRiZXN0IiwiYSI6ImNqeTQzcjMzYjE2MnMzb3AyaXBmOWZueG4ifQ.kOxbgvSal1U-3LaYKOo3rw


library(mapdeck)
mapdeck(
  # token = readLines(mapbox_token_txt),
  token = "pk.eyJ1IjoiYmRiZXN0IiwiYSI6ImNqeTQzcjMzYjE2MnMzb3AyaXBmOWZueG4ifQ.kOxbgvSal1U-3LaYKOo3rw",
  style = "mapbox://styles/bdbest/cl2139ehh000814piexs7is7f",
  zoom = 6,
  location = c(-120, 31))

# Access token: pk.eyJ1IjoiYmRiZXN0IiwiYSI6ImNqeTQzcjMzYjE2MnMzb3AyaXBmOWZueG4ifQ.kOxbgvSal1U-3LaYKOo3rw
# layer_style <- list(
#   "id"           = "hexagons",
#   "id"           = SRC_ID,
#   "type"         = "fill",
#   "source"       = "abnj_hex_res4-aysgu5",
#   "source-layer" = "country_boundaries",
#   "paint" = list(
#     "fill-color"   = "blue",
#     "fill-opacity" = 0.6))

# [
#   "interpolate",
#   ["exponential", 1],
#   ["get", "lon"],
#   -179.8177059161211,
#   "hsl(4, 82%, 42%)",
#   -0.03216,
#   "hsl(0, 0%, 96%)",
#   179.753385877314,
#   "hsl(212, 68%, 46%)"
# ]

mapboxer(
  # style = basemaps$Carto$positron,
  style = "mapbox://styles/bdbest/cl216kpw100cb15oz3ch44zui",
  center = c(0, 0),
  zoom = 1) %>%
  # add_source(vector_src, id = SRC_ID) %>%
  # add_layer(layer_style)
  add_source(mapbox_source(
    type = "vector",
    # url  = "mapbox://bdbest.1avrgtp7"), # hex_res4
    url  = "mapbox://bdbest.22bk03cc"),   # hex_res2
    id   = "src_hexagons") %>%
  add_layer(
    style = list(
      "id"           = "hexagons",
      "type"         = "fill",
      "source"       = "src_hexagons",
      # "source-layer" = "abnj_hex_res4-aysgu5", # hex_res4
      "source-layer" = "abnj_hex_res2-75qhx1",   # hex_res2
      "paint" = list(
        "fill-color"   = "blue",
        "fill-opacity" = 0.6)),
    popup = "
      <b>hexid:</b> {{hexid}}</br>
      <b>lon:</b> {{lon}}</br>
      <b>lat:</b> {{lat}}</br>
      <b>on_dtln:</b> {{on_dtln}}")
```

## mapboxapi

```{r, eval=F}
mapboxapi::tippecanoe()
mapboxapi::addMapboxTiles() # for leaflet
```


## OLD

### `st_difference()`, `st_intersection()`

```{r, eval=F}
wrld[3,] <- st_sf(
  )
hex_sf2d <- st_difference(hex_sf2, wrld[3,])

mapview(wrld[3,])
mapview(hex_sf2d)

st_intersection(wrld[3,])
hex_sf2

table(hex_sf$on_dateline, useNA = c("ifany"))

hex_sf %>% 
  filter(!on_dateline) %>% 
  mapview()

library(stringr)

i <- which(hex_sf$hexid %>% str_starts("81227"))

h <- hex_sf %>% 
  filter(!on_dateline) %>% 
  slice(1) %>% 
  pull(geometry)
m <- st_coordinates(h)[,c("X", "Y")]
d <- as.data.frame(m) %>% 
  mutate(id = 1:n())
p <- st_as_sf(d, coords = c("X", "Y"), crs=4326)
mapview(p)
st_point(m, dim="XY")
  
g <- hex_sf$geometry[i]
st_as_text(g)
# "POLYGON ((-173.2514 44.74986, -171.8731 49.05085, -177.0479 52.11352, 176.3024 50.6397, 175.621 46.18001, -179.1701 43.33462, -173.2514 44.74986))"


st_coordinates(g)[,'X']
# -173.2514 -171.8731 -177.0479  176.3024  175.6210 -179.1701 -173.2514
sort(st_coordinates(g)[,'X'])
st_coordinates(g)





g



if (hex_compact){
  hexids_c    <- h3::compact(hexids)
  glue("h3 abnj res{hex_res} compact: {length(hexids)} -> {length(hexids_c)}")
  # h3 abnj res2 compact: 2339 -> 1529
  hexids <- hex_ids_c
}

# length(hexids)

library(tidyr)


hex_sf$geom[1]

hex_sf %>% st_as_sf(crs = 4326)

%>% 
  unnest(geom) %>% 

hex_sf2 <- hex_sf %>% st_as_sf(sf_column_name = "geometry")
hex_sf2 <- hex_sf %>% st_as_sf(sf_column_name = "geometry")

class(hex_sf)
mapview(hex_sf)
h3_to_geo_boundary_sf


hex_sf <- h3_set_to_multi_polygon(hexids)
mapview(hex_sf)

x <- st_intersects(hex_sf, w)

x
```

### map `hex_ids` with `deckgl`

```{r, eval=F}
hexids_url  <- glue("https://shiny.ecoquants.com/bbnj-app/{hex}.json")
# hexids_url <- "https://shiny.ecoquants.com/bbnj-app/abnj_hex_res2.json"
# issues connecting to url so using `hexids_js` to test

# TODO: fix errors accessing json data @ associated url 
# RJSONIO::fromJSON(as.character(hexids_url))
# returns ERROR when reading from url - this seems to be why directly applying hexids_url as data source results in no h3 cluster layer on map, even though layer appears when accessing same data internally (w/o url)
  # Error in file(con, "r") : 
  # cannot open the connection to 'https://shiny.ecoquants.com/bbnj-app/abnj_hex_res3.json'
  # cannot open the connection to 'https://shiny.ecoquants.com/bbnj-app/abnj_hex_res2.json'
  # (error regardless of resolution)

properties <- list(
  stroked     = TRUE,
  filled      = TRUE,
  extruded    = FALSE,
  getHexagons = ~hexIds,
  # getFillColor = JS("d => [255, (1 - d.abnj / 500) * 255, 0]"),
  getFillColor = c(62,  136, 62),
  getLineColor = c(255, 255, 255),
  lineWidthMinPixels = 1,
  opacity = 0.4
  # ,
  # getTooltip = ~abnj
)

# with res3
deckgl(zoom = 0, pitch = 0) %>%
  # add_h3_cluster_layer(data = hexids_url, properties = properties) %>%
  add_h3_cluster_layer(data = hexids_js, properties = properties) %>%
  add_basemap()

# reading from internal json works, reading from url doesn't yet...not sure why
```

### sf example

```{r, eval=F}
# sample_data <- jsonlite::fromJSON(data_url, simplifyDataFrame = FALSE)
sf_json <- "https://raw.githubusercontent.com/uber-common/deck.gl-data/master/website/sf.h3clusters.json"

properties <- list(
  stroked = TRUE,
  filled = TRUE,
  extruded = FALSE,
  getHexagons = ~hexIds,
  getFillColor = JS("d => [255, (1 - d.mean / 500) * 255, 0]"),
  getLineColor = c(255, 255, 255),
  lineWidthMinPixels = 2,
  getTooltip = ~mean
)

deckgl(zoom = 10.5, pitch = 20) %>%
  add_h3_cluster_layer(data = sf_json, properties = properties) %>%
  add_basemap()
```

### coast

* [coordinate system - Buffer coastlines - Geographic Information Systems Stack Exchange](https://gis.stackexchange.com/questions/373691/buffer-coastlines)

```{r, eval=FALSE}
# https://gis.stackexchange.com/questions/373691/buffer-coastlines
robin = CRS("+proj=robin +lon_0=0 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")

ROI = ne_countries(returnclass = 'sf') %>%
  st_combine()

#Please check the math on this part.
#KM/earth circumference * degrees in circle
buffer_in_km <- 500
buffer_as_arc_degrees<- buffer_in_km/40075*360

coastalWaters = ROI %>%
  st_make_valid()
  st_buffer(buffer_as_arc_degrees) %>% 
   %>% 
  st_wrap_dateline()

ggplot() +
  geom_sf(data = coastalWaters, fill = "lightblue", col = "transparent") +
  geom_sf(data = ROI) +
  coord_sf(crs = robin)
```

### `h3jsr`

```{r, eval=F}
# read the shapefile of the polygon area you're interested in
nc <- st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)

# projection
nc <- st_transform(nc, crs = 4326)

# get the unique h3 ids of the hexagons intersecting your polygon at a given resolution
nc_5 <- polyfill(nc, res = 5, simple = FALSE)

# pass the h3 ids to return the hexagonal grid
hex_grid5 <- unlist(nc_5$h3_polyfillers) %>% h3_to_polygon(simple = FALSE)
hex_grid5
mapview(hex_grid5)
```


```{r}
w <- st_transform(w, crs = 4326)

st_bbox()

w <- st_sf(
  world = 1, 
  geom = st_as_sfc(
    st_bbox(
      # c(xmin = -179.999, xmax = 179.999, ymin = -86, ymax = 89.999), 
      # c(xmin = -150, xmax = 150, ymin = -50, ymax = 50), 
      c(xmin = -150, xmax = 150, ymin = -50, ymax = 50), 
      crs = st_crs(4326))))

w_p <- polyfill(w, res = 2, simple = FALSE)
length(w_p$h3_polyfillers$`0`) # -150: n=733; -180: n=371; -179: n=387

nrow(w_2)
w_2

```


