---
title: "Solutions for 5 layer scenarios"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: hide
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

Cumulative sum plots of layers for determining target per layer and allowing weights to have differential impact on scenario solution.

## Get planning units as hexagons

```{r pu}
# remotes::install_github("BenioffOceanInitiative/bbnj")
# devtools::install_local("../bbnj")
# remotes::install_github("crazycapivara/h3-r")
librarian::shelf(
  BenioffOceanInitiative/bbnj, dplyr, DT,
  exactextractr, fs, ggplot2, glue, h3, here, mapview, 
  prioritizr,
  purrr, raster, readr, rlang, scales, sf, stringr, 
  tibble,  tidyr, units)
source(here("libs/plots.R"))
select <- dplyr::select
mapviewOptions(fgb = FALSE)
mapviewOptions(
  vector.palette = mapviewPalette("mapviewSpectralColors"))
# devtools::load_all(here("~/Github/bbest/prioritizr"))
# devtools::install_local(here("~/Github/bbest/prioritizr"))

#install.packages("/Library/gurobi951/macos_universal2/R/gurobi_9.5-1_R_4.1.1.tgz")
# copied into /Library/Frameworks/R.framework/Versions/4.2/Resources/library/gurobi

hex_res <- 4 # 2
hex_geo <- glue("data/abnj_hex_res{hex_res}.geojson")
pu_rds  <- glue("data/abnj_hex_res{hex_res}_area.rds")
if (!file.exists(pu_rds)){
  
  pu <- read_sf(hex_geo)
  
  # calculate area
  pu <- pu %>%
    # st_make_valid() %>%      # dateline issues
    filter(st_is_valid(.)) %>% # 8 hexagons dropped 
    mutate(
      area_km2 = st_area(geometry) %>% 
        set_units(km^2) %>% 
        drop_units(),
      area_pct = area_km2 / sum(area_km2))
  # sum(pu$area_pct) # check adds to 12

  saveRDS(pu, file = pu_rds)
} else {
  pu <- readRDS(pu_rds)
  # paste(names(pu), collapse = ", ")
  # hexid, on_dtln, lon, lat, FID, geometry, area_km2, area_pct
}

# mapview(pu, zcol="area_pct")
# plot(pu)
```

## Get layers

```{r lyrs7}
lyrs7_grd <- here("data/lyrs7_mol.grd")

if (!file.exists(lyrs7_grd)){
  tifs1 <- list.files(here("../bbnj-scripts/map_layers"), "tif$", full.names = T)
  tifs2 <- list.files(here("../bbnj-scripts/presentation_maps"), "tif$", full.names = T)[-1] # drop redundant bio_vgpm.tif
  lyrs <- stack(c(tifs1, tifs2))
  sort(names(lyrs))
  
  s_nspp <- subset(lyrs, sort(names(lyrs))[26:48])
  r_nspp <- calc(s_nspp, sum, na.rm = T)
  
  s_mounts <- subset(lyrs, sort(names(lyrs))[49:51])
  r_mounts <- calc(s_mounts, sum, na.rm = T)
  
  r_vents <- raster(lyrs, layer = "phys_vents")
  r_scapes <- raster(lyrs, layer = "scapes_hetero")
  
  r_vgpm <- raster(lyrs, layer = "bio_vgpm")
  r_rls  <- raster(lyrs, layer = "rls_all")
  r_fishing <- raster(lyrs, layer = "fishing_KWH")
  
  # TODO: consider log/log10/normalize transforms of nspp, fish effort 
  #plot(r_vgpm)
  s_lyrs <- stack(
    r_fishing,
    r_vgpm,
    r_nspp,
    r_rls,
    r_vents,
    r_mounts,
    r_scapes)
  names(s_lyrs) <- c(
    "fishing",
    "vgpm",
    "nspp",
    "rls",
    "vents",
    "mounts",
    "scapes")
  terra::writeRaster(s_lyrs, lyrs7_grd)
}
s_lyrs7 <- terra::rast(lyrs7_grd)
  
# plot(s_lyrs7)
```

## Extract layers to planning units

```{r extract & threshold}
# extract feature values
f_0 <- exact_extract(
  s_lyrs7, pu, fun = "mean", append_cols = "hexid", progress = F) %>% 
  tibble() %>% 
  rename_with(~str_replace(., "mean.", ""), everything())

# extra calculations
f_1 <- f_0 %>% 
  mutate(
    fishing = 1/fishing,            # TODO: log(1/fishing)? 
    spp     = rls * nspp,
    benthic = vents + mounts) %>%   # TODO: rescale(vents) + rescale(mounts)? 
  mutate(
    across(where(is.numeric), ~replace_na(., 0)))
summary(f_1)

# rescale
f <- f_1 %>% 
  mutate(
    across(where(is.numeric), rescale))
summary(f)

pu_f <- pu %>%               # 129,906 × 8
  left_join(f, by = "hexid") # 129,906 × 10

get_pu_v_ks <- function(pu_f, v, ks){
  var <- as_label(enquo(v))

  d_v <- pu_f %>% 
    st_drop_geometry() %>% 
    arrange(desc({{v}})) %>% 
    mutate(
      var = var,
      km2 = cumsum(area_km2),
      val = cumsum({{v}})) %>% 
    select(var, km2, val)
  
  val_sum <- d_v %>% pull(val) %>% last()
    
  d_k <- lapply(
    ks,
    function(k){
      i <- max(which(d_v$km2 < d_v$km2[nrow(d_v)]*k))
      tibble(
        var = var, 
        k   = k,
        km2 = d_v$km2[i],
        val = d_v$val[i],
        pct_val = val/val_sum) }) %>% 
    bind_rows()

  p <- d_v %>% 
    ggplot(aes(x = km2, y = val)) + 
    geom_area(fill = "gray", color=NA)
  
  add_k_paths <- function(k, d_k){
    
    # https://colorbrewer2.org/#type=sequential&scheme=Blues&n=3
    col <- c(
      `0.3` = "#9ecae1",
      `0.5` = "#3182bd")[as.character(k)]
    
    x <- d_k %>% filter(k == !!k) %>% pull(km2)
    y <- d_k %>% filter(k == !!k) %>% pull(val)
    y_pct <- d_k %>% filter(k == !!k) %>% pull(pct_val)
    z <- tribble(
      ~x, ~y,
      x,  0,
      x,  y,
      0,  y)
    
    list(
      geom_path(data = z, aes(x,y), color=col),
      annotate(
        "text", x = x, y = y,
        label = glue("{k*100}%: {comma(x, 0.1)} ({percent(y_pct, 0.1)})")))
  }

  p <- p +
    lapply(
      ks,
      add_k_paths,
      d_k) +
    labs(
      x = "area_km2",
      y = var) +
    theme_minimal()
    
  list(
    k=d_k, p=p)
}
# kp1 <- get_pu_v_ks(pu_f, fishing, c(0.3, 0.5))
# names(k_nspp)
# k_nspp$k
# kp1$p

k_p <- sapply(
  c("fishing", "benthic", "scapes", "vgpm", "spp"),
  function(var){
    get_pu_v_ks(pu_f, !!sym(var), c(0.3, 0.5))
  }, 
  simplify = F)

p <- map(k_p, "p")

k <- map(k_p, "k") %>% 
  bind_rows() %>% 
  select(-km2) %>% 
  rename(pct_area = k)

datatable(k) %>% 
  formatPercentage(c("pct_area", "pct_val"), 1)
```

```{r plots, results = 'asis'}
# walk(p, print)
```

## Loop through scenarios

```{r}
wts <- c(0.1, 1, 10)
s_configs <- expand_grid(
  area    = c(0.3, 0.5),
  benthic = wts,
  fishing = wts,
  scapes  = wts,
  spp     = wts,
  vgpm    = wts)

s_configs %>% 
  datatable()
```

## Calculate Scenarios

```{r}
get_scenario <- function(area, benthic, fishing, scapes, spp, vgpm, redo=F, ...){
  
  o_rds <- here(glue(paste0(
      "data/scenarios/",
      "area{area}_",
      "benthic{benthic}_fishing{fishing}_scapes{scapes}_spp{spp}_vgpm{vgpm}",
      ".rds")))
  
  if (file.exists(o_rds) & !redo){
    o <- readRDS(o_rds)
  } else {
  
    # feature weights
    f_w <- c(
      benthic = benthic,
      fishing = fishing,
      scapes  = scapes,
      spp     = spp,
      vgpm    = vgpm)
    f_w <- f_w[sort(names(f_w))]
    
    # feature targets (max proportion given ocean area)
    f_t <- k %>% 
      filter(pct_area == !!area) %>% 
      select(var, pct_val) %>% 
      arrange(var) %>% 
      deframe()
    
    # problem
    p <- pu_f %>% 
      problem(features = names(f_w), cost_column = "area_pct") %>%
      add_min_shortfall_objective(area) %>% 
      add_relative_targets(f_t) %>%
      add_feature_weights(f_w)
      
    # solve
    s <- solve(p, force = T)
    
    hexids <- s %>% filter(solution_1==1) %>% pull(hexid)
    
    # calcuate actual area of solution
    # area_pct <- s %>% filter(solution_1==1) %>% pull(area_pct) %>% sum()
    
    # calculate feature representation statistics based on the prioritization
    coverage <- eval_target_coverage_summary(p, s[, "solution_1"]) %>% 
      select(feature, relative_held)
    
    o <- list(
      area_pct = area,
      weights  = f_w,
      hexids   = hexids,
      coverage = coverage)
    
    saveRDS(o, o_rds)
  }
  
  # quick fixes
  # names(o)[4] = "coverage"
  # saveRDS(o, o_rds)
  
  c(path_rds = o_rds,
    o)
}

t0 <- Sys.time()
s_lst <- s_configs %>% 
  pmap(get_scenario) # 38.3 min / 486 scenarios = 4.8 sec each
Sys.time() - t0

S <- tibble(
  s = s_lst) %>% 
  rowid_to_column("sid") %>% 
  mutate(
    scenario = map(s, function(s){
      tibble(area = s$area_pct) %>% 
        cbind(
          enframe(s$weights) %>% pivot_wider()) }),
    hexres   = hex_res,
    hexpct   = 1,
    hexid    = map(s, "hexids"),
    coverage = map(s, "coverage"))

# write all solutions to gdrive
dir_gdata <- "/Users/bbest/My Drive/projects/bbnj-app/data"

s_params <- S %>% 
  select(sid, scenario) %>% 
  unnest(scenario)
write_csv(s_params, glue("{dir_gdata}/solution_params.csv"))
write_csv(s_params, here(glue("data/solution_params.csv")))

S %>% 
  select(sid, hexres, hexid, hexpct) %>% 
  unnest(hexid) %>% 
  write_csv(glue("{dir_gdata}/solution_hexids_res{hex_res}.csv"))

S %>% 
  select(sid, coverage) %>% 
  unnest(coverage) %>% 
  write_csv(glue("{dir_gdata}/solution_coverage.csv"))

# write 10 examples for showing in Github
set.seed(42)
sid10 <- sample(S$sid, 10)
S10 <- S %>% 
  filter(sid %in% sid10)

S10 %>% 
  select(sid, scenario) %>% 
  unnest(scenario) %>% 
  write_csv(here(glue("data/sample10/solution_params.csv")))

S10 %>% 
  select(sid, hexres, hexid, hexpct) %>% 
  unnest(hexid) %>% 
  write_csv(here(glue("data/sample10/solution_hexids_res{hex_res}.csv")))

S10 %>% 
  select(sid, coverage) %>% 
  unnest(coverage) %>% 
  write_csv(here("data/sample10/solution_coverage.csv"))
```

## Aggregate to coarser hexagons and calculate percent contained

```{r}
# get_hex <- function(res=4){
#   read_sf(glue("data/abnj_hex_res{res}.geojson")) %>% 
#   filter(st_is_valid(.)) %>% 
#   mutate(
#     area_km2 = st_area(geometry) %>% 
#       set_units(km^2) %>% 
#       drop_units())
# }
# h4 <- get_hex(4)
# h3 <- get_hex(3)
# h2 <- get_hex(2)

h4 <- read_sf(glue("data/abnj_hex_res4.geojson")) %>% 
  filter(st_is_valid(.)) %>% 
  mutate(
    area_km2 = st_area(geometry) %>% 
      set_units(km^2) %>% 
      drop_units()) %>% 
  select(hexid, area_km2) %>%
  rename(
    hexid4 = hexid, 
    area4  = area_km2) %>% 
  mutate(
    hexid3 = map_chr(hexid4, h3_to_parent, 3),
    hexid2 = map_chr(hexid3, h3_to_parent, 2)) %>% 
# h0 <- h4
# h4 <- h4 %>% 
  st_drop_geometry()

h3 <- h4 %>% 
  st_drop_geometry() %>% 
  group_by(hexid3) %>% 
  summarize(
    area3 = sum(area4))

h2 <- h4 %>% 
  st_drop_geometry() %>% 
  group_by(hexid2) %>% 
  summarize(
    area2 = sum(area4))

s4_csv <- glue("{dir_gdata}/solution_hexids_res4.csv")
s3_csv <- glue("{dir_gdata}/solution_hexids_res3.csv")
s2_csv <- glue("{dir_gdata}/solution_hexids_res2.csv")

s4 <- read_csv(s4_csv) %>% 
  left_join(
    h4, 
    by = c(hexid = "hexid4"))

s3 <- s4 %>% 
  group_by(sid, hexid3) %>% 
  summarize(
    area4 = sum(area4),
    .groups = "drop") %>% 
  left_join(
    h3,
    by = "hexid3") %>% 
  mutate(
    hexres = 3,
    hexpct = area4/area3) %>% 
  select(sid, hexres, hexid=hexid3, hexpct)
write_csv(s3, s3_csv)

s3 %>% 
  filter(sid %in% sid10) %>% 
  write_csv(here(glue("data/sample10/solution_hexids_res3.csv")))

s2 <- s4 %>% 
  group_by(sid, hexid2) %>% 
  summarize(
    area4 = sum(area4),
    .groups = "drop") %>% 
  left_join(
    h2,
    by = "hexid2") %>% 
  mutate(
    hexres = 2,
    hexpct = area4/area2) %>% 
  select(sid, hexres, hexid=hexid2, hexpct)
write_csv(s2, s2_csv)

s2 %>% 
  filter(sid %in% sid10) %>% 
  write_csv(here(glue("data/sample10/solution_hexids_res2.csv")))
```

## Summary of outputs

Hi [\@mccahan](https://legendco.slack.com/team/U4HQRMYPP), I made **solutions** (n=486) for every combination of the following parameters, given by percent high seas ocean area and target conservation features (n=5):

-   **area**: 30% (0.3) or 50% (0.5)

    -   `area`: percent ocean area;

-   **features**: low (0.1), medium (1) or high (10)

    -   `benthic`: benthic features: seamounts + hydrothermal vents

    -   `fishing`: inverse of kilowatt hours fished; i.e. avoid highly fished areas

    -   `scapes`: seascapes; i.e. heterogeneity of seafloor

    -   `spp`: species importance given by species richness \* species extinction risk

    -   `vgpm`: primary productivity, given by the Vertically Generalized Production Model

Outputs can be found here in [bbnj-app/data/ - Google Drive](https://drive.google.com/drive/u/0/folders/1dbobDmaxZs4T9wUcFnZ3AboNXcEusDdU):\

-   spatial hexagons unique by `hexid`:

    -   `hex_res4.geojson`(88 mb): resolution 4 (\~ 41 km width)

    -   `hex_res3.geojson`(33 mb): resolution 3 (\~ 107 km width)

    -   `hex_res2.geojson`(24 mb): resolution 2 (\~ 266 km width)

-   input parameters:

    -   `solution_params.csv`(11 kb): unique parameter values of area and features, identified by integer scenario id (`sid`)

-   output hexagons (`hexid`)  included in each scenario (`sid`):

    -   `solution_hexids_res4.csv`(605 mb): finest resolution that acted as fundamental planning unit in prioritization, so percentage always 100% (`hexpct`: 1)

    -   `solution_hexids_res3.csv`(179 mb): medium resolution hexagon aggregated to see when zoomed out (`hexpct`: ≤1)

    -   `solution_hexids_res2.csv`(45 mb): coarsest resolution of aggregation for when fully zoomed out (`hexpct`: ≤1)

-   feature coverage:

    -   `solution_coverage.csv`(69 kb): percent held per feature (n=5) for each scenario (`sid`)

I'll work on a prototype app to crudely display these results from slider inputs, then work on varying subtaxa within species.
